---
title: "Reactive Shiny Apps"
author: "Prakhar"
date: '2023-09-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plotly)
library(shiny)
library(miniUI)
library(dplyr)
library(readr)
library(tidytext)
library(visNetwork)
source("phraseNet.R")
```

## Introducing trace line colors in plotly.js parallel coordinates plot

Parallel coordinate plots in plotly provide the functionality to select trace lines by brushing along the parallel coordinate axes. However all the selected clusters have the same color making it difficult to differentiate between clusters. The following Shiny gadget makes use of reactivity in Shiny to enable color selection prior to brusing the trace lines, and makes those colors persist in the subsequent brushes making it much more easier to identify different clusters in the data. We can also remove the colors of the clusters in the reverse order of cluster selection to account for erroneous brushing by the user or to enable corrections in the cluster selection.
```{r}
# special Task 2
brush_trace_lines <- function(){
  # create user interface with one plot and two buttons
  ui <- miniPage(gadgetTitleBar("Brush Trace Lines in the Parallel Coordinates Plot of mtcars dataset with the selected colors"),
                 miniContentPanel(
                   selectInput(inputId = "picked_color",
                               label = "pick a color",
                               choices = list(red = 1, green = 2, blue = 3, orange = 4)),
                   actionButton(inputId = "clear_sel", label = "Clear Selected"),
                   actionButton(inputId = "clear_all", label = "Clear All"),
                   plotlyOutput(outputId = "parcoords"),
                 )
  )
  
  server <- function(input, output) {
    # use mtcars data and set ID column
    mtcars2 <- mtcars
    # mtcars2$ID <- 1:nrow(mtcars)
    # sequence of colors selected by the user
    color_seq <- reactiveVal(0)
    # color of individual observations
    brushed_colors <- reactiveVal(rep(0, nrow(mtcars)))
    # observe the event for clear selected button and react accordingly
    observeEvent(input$clear_sel, {
      n_colors <- length(color_seq())
      cur_color <- color_seq()[n_colors]
      bc <- brushed_colors()
      # grey out the current color
      bc[bc == cur_color] <- 0
      # update the brushed colors
      brushed_colors(bc)
      # update the color sequence
      if(n_colors > 1)
        color_seq(color_seq()[1:(n_colors-1)])
    })
    # observe the event for clear all button and react accordingly
    observeEvent(input$clear_all, {
      # update the color sequence
      color_seq(0)
      # update the brushed colors
      brushed_colors(rep(0, nrow(mtcars)))
    })
    # render the parallel coordinates plot and register the restyle event
    output$parcoords <- renderPlotly({
      # create dimensions list
      dims <- Map(function(x, y) {
        list(values = x, range = range(x), label = y)
      }, mtcars2, names(mtcars2), USE.NAMES = FALSE)
      # plot the parallel coordinates
      plot_ly(type = 'parcoords', dimensions = dims[1:11], source = "pcoords",
              line = list(color = ~brushed_colors(),
                          colorscale = list(c(0, "grey"),
                                            c(0.25, "red"),
                                            c(0.5, "green"),
                                            c(0.75, "blue"),
                                            c(1, "orange")),
                          cmin = 0,
                          cmax = 4
              )
      ) %>% 
        layout(margin = list(r = 30)) %>%
        event_register("plotly_restyle")
    })
    # observe the brushing event and redraw the plot with selected color
    observeEvent(event_data("plotly_restyle", source = "pcoords"), {
      d <- event_data("plotly_restyle", source = "pcoords")
      # what is the relevant dimension (i.e. variable)?
      dimen <- as.numeric(stringr::str_extract(names(d[[1]]), "[0-9]+"))
      # If the restyle isn't related to a dimension, exit early.
      if (!length(dimen)) return()
      # careful of the indexing in JS (0) versus R (1)!
      dim_name <- names(mtcars2)[[dimen + 1]]
      # get the constraint range
      rng <- d[[1]][[1]]
      # get the IDs from the dimension and constraint range
      IDs <- which(between(mtcars2[[dim_name]], rng[1], rng[2]))
      # use the IDs to change the brushed colors and the color sequence
      color_seq(c(color_seq()[color_seq() != input$picked_color], input$picked_color))
      bc <- brushed_colors()
      bc[IDs] <- input$picked_color
      brushed_colors(bc)
    })
    # observe the done button to stop the gadget and return the colors of each 
    # observation
    observeEvent(input$done, {
      color <- as.numeric(brushed_colors()) + 1
      color <- sapply(color, function(x) {
        switch(x,
               "grey",
               "red",
               "green",
               "blue",
               "orange")
      })
      mtcars2$color <- color
      # print the data frame with colors
      print(mtcars2)
      # return the data frame with colors
      stopApp(mtcars2)
    })
  }

  runGadget(ui, server)
}
# run the gadget
Result <- brush_trace_lines()
```

## Phrase Networks

This reactive Shiny app reads any user input text file and takes a list of connector words, then displays a graph of all the words in the texts connected by those connector words referred to as a phrase net. Essentially displaying all the trigrams in the text in which the middle word is one of the connector words input by the user. The nodes are also colored by their occurrence frequency and the positions in the trigrams, with words most commonly occurring as the first word (source node) in the trigram colored in darker shades of blue and the one most commonly occurring as the last word (destination node) colored in lighter shades. This helps us get an overview of the text by analyzing the different connections in the text and their relative frequencies. 
Further, selecting any edge of the graph displays all the sentences in which the connection appears, which helps to better understand the context in which the phrases are used to get a better understanding of the text.

A visually more impressive version of this app is also hosted as a shiny dashboard here: https://89lvi7-prakhar.shinyapps.io/PhraseNetworkApp/
```{r}
# Special Task 3
ui <- fluidPage(
  
  sidebarLayout(
    
    sidebarPanel(
      # select a file to read
      fileInput(inputId = "file", label = "select a text file", accept = "text/plain"),
      # input the connectors
      textInput(inputId = "connectors", label = "Enter comma separated list of connectors")
    ),
    
    mainPanel(
      # output the phrase net
      visNetworkOutput(outputId = "phrasenet"),
      # text output for the sentences containing the selected edge
      htmlOutput(outputId = "text")
    )
  )
)

server <- function(input, output) {
  
  text <- reactiveVal()
  # observe the file input and change the text accordingly
  observeEvent(input$file, {
    text(read_lines(input$file$datapath))
  })
  
  connectors <- reactiveVal()
  # observe the connectors input and change the connectors accordingly
  observeEvent(input$connectors, {
    connector_text <- input$connectors
    # vector of connector words
    connector_words <- strsplit(connector_text, ",")[[1]]
    # clean the connector words by stripping the white space
    connector_words <- gsub(" ", "", connector_words)
    connectors(connector_words)
  })
  
  # edge data obtained from phraseNet function
  edge_data <- reactiveVal()
  
  # output the phrasenet
  output$phrasenet <- renderVisNetwork({
    # get the nodes and edges for the graph, using the text and connectors
    graph_data <- phraseNet(text(), connectors())
    # change the edge data for use in text output
    edge_data(graph_data$edges)
    # draw the graph and associate a click event with the edges
    visNetwork(graph_data$nodes, graph_data$edges) %>%
      visEvents(selectEdge = "function(edges) {
        Shiny.onInputChange('current_edge_id', edges);
      ;}")
  })
  
  # output the relevant lines using the text output
  output$text <- renderUI({
    # get the id of the selected edge
    edge_id <- input$current_edge_id$edges[[1]]
    # get the from and to words of the selected edge
    edge <- edge_data() %>% 
      filter(id == edge_id)
    from <- edge$from
    to <- edge$to
    # get the word combinations using the from, to and connectors
    word_combinations <- paste(from, connectors(), to)
    # get the output lines for each word combination
    HTML(paste(unlist(sapply(word_combinations, grep, x = text(), ignore.case = TRUE, value = TRUE)), collapse = "<br>"))
  })
}

# Run the application 
shinyApp(ui, server)
```

## Acknowledgements

The code for creating the phrase network is taken and modified from the R script "phraseNet.R" given in the LiU course 732A98 webpage: https://www.ida.liu.se/~732A98/info/materials.en.shtml